---
title: "sample_from_density()"
output:
  html_document: default
  pdf_document: default
vignette: |
  %\VignetteIndexEntry{read_med() vignette} 
  %\VignetteEngine{knitr::rmarkdown} 
  \usepackage[utf8]{inputenc}
---

```{r echo = FALSE, warning=FALSE}
library(YEAB)
```

## Introduction

Behavioral data often exhibit complex patterns that may not adhere to well-known parametric distributions. Traditional statistical methods reliant on such distributions may therefore prove inadequate in capturing the nuanced behaviors observed in real-world scenarios.

Approximating a density distribution from a data sample using Kernel Density Estimation allows researchers to estimate the underlying patterns that could emerge from a specific experiment, thus, the ability to generate synthetic data samples that closely mimic the distribution of observed behavioral data serves as a valuable tool in behavioral analysis. Such samples allow researchers to explore various hypotheses, validate statistical models, and assess the efficacy of experimental interventions in a controlled setting.

Here we introduce the `sample_from_density()` function which generates a given `n` amount of data points from a density distribution calculated using KDE (Kernel Density Estimation) from a given data set.

The function takes two parameters:

- `x` A numeric vector of data points from an distribution.
- `n` the number of samples to return.

## Example

Let's generate a random sample from a normal distribution:
```{r, echo = FALSE, warning= FALSE, message = FALSE}
library(foreach)
library(doParallel)
library(parallel)
no_cores <- detectCores()
registerDoParallel(makePSOCKcluster(no_cores))
```

```{r}
set.seed(420)
normal_sample <- rnorm(100)
```

Now let's create a sample of 100 data points from the distribution estimated with the `sample_from_density()` function:

```{r}
sampled_dist <- sample_from_density(normal_sample, 100)
```

Finally let's compare both distributions:
```{r, echo = FALSE}

plot(density(normal_sample), xlab = "", main = "rnorm() vs sample_from_density()",
     col = "blue")
lines(density(sampled_dist), col = "red")
legend("topleft", c("sample_from_density", "rnorm"), 
       lty = c(1, 1), col= c("red", "blue"))

```

```{r, eval = FALSE, echo = FALSE}

res <- foreach(1:10000, .combine = c,
                .packages=c("infotheo", "YEAB"))  %dopar% {
    x2 <- sample_from_density(x, 100)
    na.omit(x2)
    KL_div(x, x2, min(c(x, x2)), max(c(x, x2)))
                }

 hist(res)
```